---
title: "herringr-carbonate"
author: "Kate Morkeski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Libraries used

```{r}

library(here)
#library(readxl)
library(tidyverse)

```

Read in carbonate data

```{r}

dicta <- read_csv(here('herringriv_carbonate_summary.csv'), show_col_types = FALSE) 

dicta$Date <- as.Date(dicta$Date, format = "%m/%d/%Y")  

dicta$Treatment <- as.factor(dicta$Treatment)
dicta$Depth_cm <- as.factor(dicta$Depth_cm)

```

Organize data

```{r}

dicta$Sample_ID <- gsub("22-", "", dicta$Sample_ID)
dicta$Sample_ID <- as.integer(dicta$Sample_ID)
dicta <- dicta[order(dicta$Sample_ID),]

dicta$Month <- format(dicta$Date, "%m")

dicta <- dicta %>% mutate(Spur = case_when(Plot > 7 ~ "creek",
                                       Plot < 8 ~ "woods"))

```
Normalize by salinity

```{r}
#TODO: change placeholder salinity value to actual creek salinity for this date or tide
dicta <- dicta %>% mutate(DIC_norm = DIC_umol_kg*22/Salinity_psu)
dicta <- dicta %>% mutate(TA_norm = TA_umol_kg*22/Salinity_psu)

```

Break out data tables by month and depth

```{r}

dicta_Aug <- dicta %>% filter(Month == '08')

dicta_Aug_5 <- dicta_Aug %>% filter(Depth_cm == 5)
dicta_Aug_15 <- dicta_Aug %>% filter(Depth_cm == 15)
dicta_Aug_sand <- dicta_Aug %>% filter(Depth_cm == 3)

dicta_15 <- dicta %>% filter(Depth_cm == 15)

ggplot(dicta, aes(TA_umol_kg, DIC_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title= "All measured DIC and TA by treatment")
ggplot(dicta, aes(TA_umol_kg, DIC_umol_kg, color = Month)) + geom_point(size = 4) + labs(title= "All measured DIC and TA by month")
ggplot(dicta, aes(TA_umol_kg, DIC_umol_kg, color = Depth_cm)) + geom_point(size = 4) + labs(title= "All measured DIC and TA by sampling depth")

```




# Plot August 5 cm data

```{r}

#trt_colors <- c()

ggplot(dicta_Aug_5, aes(Salinity_psu, DIC_umol_kg, color = Spur)) + geom_point(size = 4)+ labs(title = "T5 August 5cm DIC vs salinity, by spur")
ggplot(dicta_Aug_5, aes(Salinity_psu, DIC_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title = "T5 August 5cm DIC vs salinity, by treatment")

ggplot(dicta_Aug_5, aes(Salinity_psu, TA_umol_kg, color = Spur)) + geom_point(size = 4)+ labs(title = "T5 August 5cm TA vs salinity, by spur")
ggplot(dicta_Aug_5, aes(Salinity_psu, TA_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title = "T5 August 5cm TA vs salinity, by treatment")

ggplot(dicta_Aug_5, aes(Treatment, Salinity_psu)) + geom_boxplot()  + labs(title = "T5 August 5cm salinity")

ggplot(dicta_Aug_5, aes(Treatment, DIC_umol_kg)) + geom_boxplot() + labs(title = "T5 August 5cm DIC")

ggplot(dicta_Aug_5, aes(Treatment, TA_umol_kg)) + geom_boxplot() + labs(title = "T5 August 5cm TA")

ggplot(dicta_Aug_5, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T5 August 5cm pH")

ggplot(dicta_Aug_5, aes(TA_umol_kg, DIC_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title = "T5 August 5cm DIC vs TA")

ggplot(dicta_Aug_5, aes(Treatment, DIC_norm)) + geom_boxplot()+ labs(title = "T5 August 5cm DIC normalized with placeholder salinity = 22 psu")

ggplot(dicta_Aug_5, aes(Treatment, TA_norm)) + geom_boxplot()+ labs(title = "T5 August 5cm TA normalized with placeholder salinity = 22 psu")


```
Plot normalized DIC and TA without plot 4, which has higher salinity and lower DIC and TA than other quartz plots. 

```{r}

dicta_Aug_5_no_plot_4 <- dicta_Aug_5 %>% filter(Plot != 4)

ggplot(dicta_Aug_5_no_plot_4, aes(Treatment, DIC_norm)) + geom_boxplot()+ labs(title = "T5 August 5cm DIC normalized with placeholder salinity = 22 psu" , subtitle = "excluding quartz plot #4")

ggplot(dicta_Aug_5_no_plot_4, aes(Treatment, TA_norm)) + geom_boxplot()+ labs(title = "T5 August 5cm TA normalized with placeholder salinity = 22 psu" , subtitle = "excluding quartz plot #4")

```


Plot sand layer data

```{r}

ggplot(dicta_Aug_sand, aes(Salinity_psu, DIC_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title = "T5 August sand layer DIC vs salinity")
ggplot(dicta_Aug_sand, aes(Salinity_psu, TA_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title = "T5 August sand layer TA vs salinity")

ggplot(dicta_Aug_sand, aes(Treatment, Salinity_psu)) + geom_boxplot()  + labs(title = "T5 August sand layer salinity")

ggplot(dicta_Aug_sand, aes(Treatment, DIC_umol_kg)) + geom_boxplot()+ labs(title = "T5 August sand layer DIC")
ggplot(dicta_Aug_sand, aes(Treatment, TA_umol_kg)) + geom_boxplot()+ labs(title = "T5 August sand layer TA")

ggplot(dicta_Aug_sand, aes(Treatment, pH)) + geom_boxplot()+ labs(title = "T5 August sand layer pH")

```
# Plot August 15 cm data

```{r}

ggplot(dicta_Aug_15, aes(Salinity_psu, DIC_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title = "T5 August 15cm")
ggplot(dicta_15, aes(Salinity_psu, DIC_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title = "All measured 15cm")
ggplot(dicta_15, aes(Salinity_psu, DIC_umol_kg, color = Spur)) + geom_point(size = 4)+ labs(title = "All measured 15cm")
ggplot(dicta_15, aes(Salinity_psu, DIC_umol_kg, color = Month)) + geom_point(size = 4)+ labs(title = "All measured 15cm")

ggplot(dicta_Aug_15, aes(Salinity_psu, TA_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title = "T5 August 15cm")
ggplot(dicta_15, aes(Salinity_psu, TA_umol_kg, color = Treatment)) + geom_point(size = 4)+ labs(title = "All measured 15cm")
ggplot(dicta_15, aes(Salinity_psu, TA_umol_kg, color = Spur)) + geom_point(size = 4)+ labs(title = "All measured 15cm")
ggplot(dicta_15, aes(Salinity_psu, TA_umol_kg, color = Month)) + geom_point(size = 4)+ labs(title = "All measured 15cm")

```
Read in sensor data

```{r}

sensor <- read_csv(here('porewater_sensor_data.csv'), show_col_types = FALSE) 

sensor$Date <- as.Date(sensor$Date, format = "%m/%d/%Y")  

sensor$Treatment <- as.factor(sensor$Treatment)

sensor$Month <- format(sensor$Date, "%m")

sensor <- sensor %>% mutate(Spur = case_when(Plot > 7 ~ "creek",
                                       Plot < 8 ~ "woods"))

```

Create dataframe for each timepoint

```{r}

sensor_T0_5cm <- sensor %>% filter(Timepoint == 'T0' & Tip_Depth_cm == "5")
sensor_T0_15cm <- sensor %>% filter(Timepoint == 'T0' & Tip_Depth_cm == "15")

sensor_T1_5cm <- sensor %>% filter(Timepoint == 'T1' & Tip_Depth_cm == "5")
sensor_T1_15cm <- sensor %>% filter(Timepoint == 'T1' & Tip_Depth_cm == "15")

sensor_T2_5cm <- sensor %>% filter(Timepoint == 'T2' & Tip_Depth_cm == "5")
sensor_T2_15cm <- sensor %>% filter(Timepoint == 'T2' & Tip_Depth_cm == "15")

sensor_T3_5cm <- sensor %>% filter(Timepoint == 'T3' & Tip_Depth_cm == "5")
sensor_T3_15cm <- sensor %>% filter(Timepoint == 'T3' & Tip_Depth_cm == "15")
sensor_T3_sand <- sensor %>% filter(Timepoint == 'T3' & Tip_Depth_cm != "5" & Tip_Depth_cm != "15")

sensor_T4_5cm <- sensor %>% filter(Timepoint == 'T4' & Tip_Depth_cm == "5")
sensor_T4_15cm <- sensor %>% filter(Timepoint == 'T4' & Tip_Depth_cm == "15")
sensor_T4_sand <- sensor %>% filter(Timepoint == 'T4' & Tip_Depth_cm != "5" & Tip_Depth_cm != "15")

sensor_T5_5cm <- sensor %>% filter(Timepoint == 'T5' & Tip_Depth_cm == "5")
sensor_T5_15cm <- sensor %>% filter(Timepoint == 'T5' & Tip_Depth_cm == "15")
sensor_T5_sand <- sensor %>% filter(Timepoint == 'T5' & Tip_Depth_cm != "5" & Tip_Depth_cm != "15")

sensor_T6_5cm <- sensor %>% filter(Timepoint == 'T6' & Tip_Depth_cm == "5")
sensor_T6_15cm <- sensor %>% filter(Timepoint == 'T6' & Tip_Depth_cm == "15")
sensor_T6_sand <- sensor %>% filter(Timepoint == 'T6' & Tip_Depth_cm != "5" & Tip_Depth_cm != "15")

# sensor_T1 <- sensor %>% filter(Timepoint == 'T1')
# sensor_T2 <- sensor %>% filter(Timepoint == 'T2')
# sensor_T3 <- sensor %>% filter(Timepoint == 'T3')
# sensor_T4 <- sensor %>% filter(Timepoint == 'T4')
# sensor_T5 <- sensor %>% filter(Timepoint == 'T5')
# sensor_T6 <- sensor %>% filter(Timepoint == 'T6')

```
# Plot sensor data

5 cm salinity

```{r}

ggplot(sensor_T0_5cm, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T0 May 5 cm salinity")
ggplot(sensor_T1_5cm, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T1 May 5 cm salinity")
ggplot(sensor_T2_5cm, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T2 May 5 cm salinity")
ggplot(sensor_T3_5cm, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T3 June 5 cm salinity")
ggplot(sensor_T4_5cm, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T4 July 5 cm salinity")
ggplot(sensor_T5_5cm, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T5 August 5 cm salinity")
ggplot(sensor_T6_5cm, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T6 September 5 cm salinity")

```
# 5 cm pH
```{r}

ggplot(sensor_T0_5cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T0 May 5 cm pH")
ggplot(sensor_T1_5cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T1 May 5 cm pH")
ggplot(sensor_T3_5cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T2 May 5 cm pH")
ggplot(sensor_T4_5cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T3 June 5 cm pH")
ggplot(sensor_T5_5cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T5 August 5 cm pH")
ggplot(sensor_T6_5cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T6 September 5 cm pH")

```
Sand salinity

```{r}

ggplot(sensor_T3_sand, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T3 June sand layer salinity")
ggplot(sensor_T4_sand, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T4 July sand layer salinity")
ggplot(sensor_T5_sand, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T5 August sand layer salinity")
ggplot(sensor_T6_sand, aes(Treatment, sal_psu)) + geom_boxplot() + labs(title = "T6 September sand layer salinity")

```
Sand pH

```{r}

#ggplot(sensor_T3_sand, aes(Treatment, pH)) + geom_boxplot()  # no pH data
ggplot(sensor_T4_sand, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T4 July sand layer pH")
ggplot(sensor_T5_sand, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T5 August sand layer pH")
ggplot(sensor_T6_sand, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T6 September sand layer pH")

```
15 cm salinity

```{r}

ggplot(sensor_T0_15cm, aes(Treatment, sal_psu)) + geom_boxplot()+ labs(title = "T0 May 15 cm salinity")
ggplot(sensor_T1_15cm, aes(Treatment, sal_psu)) + geom_boxplot()+ labs(title = "T1 May 15 cm salinity")
ggplot(sensor_T2_15cm, aes(Treatment, sal_psu)) + geom_boxplot()+ labs(title = "T2 May 15 cm  salinity")
ggplot(sensor_T3_15cm, aes(Treatment, sal_psu)) + geom_boxplot()+ labs(title = "T3 June 15 cm salinity")
ggplot(sensor_T4_15cm, aes(Treatment, sal_psu)) + geom_boxplot()+ labs(title = "T4 July 15 cm salinity")
ggplot(sensor_T5_15cm, aes(Treatment, sal_psu)) + geom_boxplot()+ labs(title = "T5 August 15 cm salinity")
ggplot(sensor_T6_15cm, aes(Treatment, sal_psu)) + geom_boxplot()+ labs(title = "T6 September 15 cm salinity")

```

15 cm pH

```{r}

ggplot(sensor_T0_15cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T0 May 15 cm pH")
ggplot(sensor_T1_15cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T1 May 15 cm pH")
ggplot(sensor_T2_15cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T2 May 15 cm pH")
ggplot(sensor_T3_15cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T3 June 15 cm pH")
ggplot(sensor_T4_15cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T4 July 15 cm pH")
ggplot(sensor_T5_15cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T5 August 15 cm pH")
ggplot(sensor_T6_15cm, aes(Treatment, pH)) + geom_boxplot() + labs(title = "T6 September 15 cm pH")

```
